The Spatial and Temporal Domains of Modern Ecology

### 2018-02-03 paper/main.tex
Final edits for re-submission

%\linenumbers
\vspace{10pt}
\noindent \textbf{Introduction}
\vspace{5pt}
\\
The scales at which ecosystems are observed play a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.  Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is essential for understanding ecological phenomena \cite{levin_problem_1992,sandel_scale_2009}. This awareness has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad and increasing range of scales \cite{schneider_rise_2001, chave_problem_2013}.

Given the growing awareness of scale, expanding data gathering capabilities, and the fact that the most comprehensive (and arguably best-known) meta-analyses \cite{tilman_ecological_1989,kareiva_spatial_1988} of ecological research scales were published nearly 30 years ago (but see \cite{porter_crop_2005,sandel_scale_2009} for more recent reviews), it is both timely and important to assess the scales of contemporary ecological investigation. To address this need, we quantified the spatial and temporal domains of empirical observations that were reported within recently (2004-2014) published ecological studies. We define domain as the distribution of observations within the spectrum of one or more scale dimensions\footnote{This definition differs slightly from Wiens' \cite{wiens_spatial_1989}, who defined ``domain of scale'' as ``a portion of the scale spectrum within which process-pattern relationships are consistent regardless of scale.''}, and empirical observations as ecological observations collected under un-controlled or non-manipulated conditions. Empirical observations are critical for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992, tilman_ecological_1989}, therefore the spatio-temporal domains of observations provide an important indicator of the field's progress towards achieving a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.

@@

\vspace{10pt}
\noindent \textbf{Results}
\vspace{5pt}
\\
We reviewed 348 papers randomly selected from 42,918 titles published between 2004-2014 in the top 30 ecology-themed journals. We extracted scale data from 378 observations of ``natural'' (i.e. non-experimentally manipulated) ecological features reported within 133 of the reviewed papers (plus an additional 62 cited as the source of observations). The majority of sampled observations were collected using conventional field methods (80\%), followed by automated \emph{in situ} sensing techniques (12.4\%), remote sensing (6.9\%), and paleo-reconstruction ($<$0.8\%).

\vspace{5pt}
\noindent \textbf{\emph{Observational domains within individual dimensions}}\\
In terms of resolution, spatial replicates for the majority (67\%) of observations (across all methods) were $\leq$1 m$^2$, a further 24\% were 1 m$^2$ up to 1 ha, and the remaining 9\% were $>$1 ha (Fig. 1A). These distributions primarily reflect those of field observations, the dominant observational methodology. Examining the distributions for each observational method (Supplementary Figure 1) shows that automated sensing and paleo-reconstruction had resolutions that were generally finer (85\% or more $\leq$0.1 m$^2$) than field observations (47\% $\leq$0.1 m$^2$), while the majority of remote observations were much coarser (70\% $>$100 m$^2$).

The extent of 19\% of observations was $\leq$10 ha, 23\% covered 10-1,000 ha, 11\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 15\% $>$1,000,000 ha (Fig. 1B).  As with resolution, the extent covered by automated sensing methods tended to be smaller (52\% $\leq$100 ha) than those of field observations (31\% $<$100 ha), while all but 4\% of remote observations covered areas $>$10,000 ha (as did the small number of paleo-reconstructions).

%16 + 29 + 26 + 67 + 14 + 37 + 10

@@

Duration was $\leq$1 day for 31\% of sampled observations (due to lack of temporal replication), while 10\% covered one day to one month, 23\% one month to one year, 27\% 1-10 years, and 9\% a decade or more (several paleoecological studies covered centuries; Fig. 1D). Paleo-reconstructions naturally had the longest duration (67\% $>$ 1 decade), while just $\sim$40\% of field, automated, and remote observations had durations exceeding one year.

\vspace{5pt}
\noindent \textbf{\emph{Observational domains within two-dimensions}}\\
Contrasting resolution with interval reveals that the majority of temporally replicated observations (unrepeated observations were excluded because they lack intervals) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig. 2A). A less dense, oblong concentration of observations bounded on the upper left by monthly to yearly observations at 100 m$^2$ resolution and on the lower right by near-daily to monthly observations with 1-10 ha resolution is also evident. The four observational methods occupied substantially different domains, as indicated by the locations of their median values (and see Supplementary Figure 2): the median domain of field observations had 0.1-1 m$^2$ resolution and a monthly interval, whereas remote observations had a coarser median resolution (1,000 m$^2$) but finer median interval ($\sim$1 day). Paleo-reconstructions and automated sensing techniques were both finely resolved (median between 10 cm$^2$-0.01 m$^2$), but automated approaches had an hourly-daily median interval compared to multi-decadal for paleo-reconstructions.

Comparing the interval and duration of temporally replicated observations showed most observations had daily to decadal intervals and durations of 1 month to 1 decade (Fig. 2B). The orientation of this concentration shows that interval increases with duration; observations lasting one month to one year tend to have daily to monthly intervals, while those lasting 1 year to 1 decade tend to have yearly to decadal intervals. This tendency is reflected in the domain medians of the primary observational methods: automated sensing had the finest median interval (hour-day) and shortest duration (month-year), followed by remote sensing (median interval slightly greater than one day and median duration of 1 year), field observations (median monthly interval and duration just over 1 year), and finally paleo-reconstructions (median interval 1 decade and millennial duration).

Contrasting the two spatial dimensions against one another shows a primary concentration of observations of 10 cm$^2$ to nearly 100 m$^2$ resolution with extents ranging between $\sim$1,000-1,000,000 ha (Fig. 2C). The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution and 1000 m$^2$ to $\leq$10 ha extent. These three concentrations suggest that observational extent increases with resolution, which is further evident in the median domain values (and kernel densities; Supplementary Figure 2) of automated (0.01 m$^2$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.  Paleo-reconstructions were an outlier from this relationship, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a possible artifact of small sample size.

There are two primary observational domains within the contrast between duration and extent. The first consists of observations lasting one month to one decade with extents of 10-1,000 ha, while the second is defined by observations of one year to several decades that cover 10,000-1,000,000 ha (Fig. 2D).  Three other notable, but lesser concentrations are also evident, including small area observations (0.1-1 ha) covering one month to decade, and short duration, temporally unreplicated observations ($\leq$1 day) of either 1-100 ha or 10,000-1,000,000 ha.  The median observation from automated sensing (1 year duration, 100 ha extent) lies near the center of the first major concentration, while the median extents of field (1,000-10,000 ha) and remote (1-10 million ha) observations bound the second major concentration at its upper and lower extents, with the median duration of both observational types falling between 1 month to 1 year.

\vspace{5pt}
\noindent \textbf{\emph{Differences between actual and ostensible scales}}\\
Observational extent was on average 5.6 orders of magnitude larger than actual extent (Fig. 3A). This difference increased with extent, reaching a maximum of 8.3 between 100 million and 1 billion ha of extent, then falling to 3 orders of magnitude between 1-10 billion ha (these extents were covered by $<$2\% of observations, which were primarily collected with remote sensing). Remote observations had the smallest mean difference magnitude (1.9), compared to 5.7 or larger for the other three methods (Supplementary Figure 3).

The magnitudes of difference between observational duration and actual duration was somewhat smaller, averaging 3.4 and ranging from just under 2 for the shortest durations (hour-day) to over 4 for observations lasting 1 decade to 1 century (Fig. 3B). As with extent, the difference fell substantially for the longest durations (century to millenia), as these domains were covered by paleo-reconstructions (Supplementary Figure 3), which have little difference between actual and ostensible duration because coring techniques capture continuous temporal records. The mean difference magnitudes for the other three observing methods ranged from just over 3 (field and automated sensing) to nearly 6 (remote observations).

\vspace{5pt}
\noindent \textbf{\emph{Potential biases and uncertainties in quantifying scales}}\\
Our results were potentially influenced by several methodological issues. First, most studies did not precisely report observational scales, thus we had to estimate, rather than simply record, scale values for the majority of observations (we estimated 63, 60, 69, 36, 64, and 83\% of resolution, extent, actual extent, interval, duration, and actual duration values, respectively). Estimation errors may therefore have biased our findings. We attempted to quantify and account for this error by assessing between-observer variability and incorporating this uncertainty into our resampling methodology (Supplementary Results). The resulting confidence intervals (Fig. 1) suggest that that estimation errors did not unduly influence our findings.

Our scale-estimation protocols may also have introduced bias, particularly our rule for estimating resolution (the smallest areal unit of \emph{complete} measurement). We selected this definition for the sake of consistency, but some papers reported resolution as a larger area in which sub-samples were taken. For these, our estimates were finer than what the studies' authors considered to be the resolution. Our results would also  be somewhat different if we had included observations from experiments. For example, average resolution and duration would likely be finer \cite{tilman_ecological_1989,kareiva_spatial_1988}. Additionally, the token 1 second (Supplemental Methods) we used to represent the duration of remotely sensed temporal replicates (which are effectively instantaneous) caused us to underestimate the differences between their durations and actual durations (Supplementary Figure 3). However, the relatively small number of remote observations suggests that the impact of this bias on our overall findings was negligible.

@@

Finally, our omission of papers published after 2014 could also have biased our findings. Although our sample size was too small to assign statistical significance, we found a possible positive trend in the use of remote observations and a corresponding decline in field observations over the course of our study period. If these trends were not spurious, they suggest that including studies from 2015-2017 would result in a somewhat larger relative sample of remote observations, which would in turn slightly increase the mean extent of ecological observations (see Supplementary Results for further details).

\vspace{10pt}
\noindent \textbf{Discussion}
\vspace{5pt}
\\
\noindent \textbf{\emph{Insights into the scale domains of modern ecology}}\\
Our results suggest that modern ecology's observational domains are fairly narrow, and that ecologists still primarily rely on conventional field-based observational techniques. In the spatial dimensions, the majority of observations have grains of $\leq$1 m$^2$ and extents of $\leq$10,000 ha (Fig 1A;B). In the temporal dimensions, most observations are either un-replicated or relatively infrequent ($>$1 month interval; Fig. 1C) and have relatively short duration ($\leq$1 year; Fig. 1D).

Contrasting observational dimensions reveals that larger extents are associated with larger spatial replicates (Fig. 2C), while longer durations are associated with longer intervals (Fig. 2B). The latter association reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of field observations, but also appears to affect the other three methods, as evidenced by their relative domain locations. A similar tradeoff can be seen in the inverse relationship between resolution and interval (Fig. 2A), which primarily relates to field observations, where larger spatial replicates demand greater effort, which in turn reduces sampling frequency \cite{kareiva_spatial_1988}. Less obvious is the opposite tradeoff that affects remote observation (Supplementary Figure 2), where finer resolution (which is necessary for detail) typically necessitates longer intervals \cite{estes_platform_2016}.

As a result of these tradeoffs, there are several notable observational gaps, specifically within the domains defined by high frequency (daily to sub-daily intervals) observations having 1) high to moderate resolutions ($>$1 m$^2$ up to 100 ha; Fig. 2A) and 2) decadal or longer durations (Fig. 2B).  Another hole is evident in the high to moderate resolution, large extent (1 million-10 billion ha) domain (Fig. 2C).

Have these domains changed since the seminal papers on scale first appeared in the late 1980s \cite{wiens_spatial_1989, levin_problem_1992, tilman_ecological_1989}? A comprehensive answer would require a repeat similar analysis focused on earlier literature, but the data provided by three prior studies provides partial insight.  The first dataset consists of duration values that Tilman \cite{tilman_ecological_1989} extracted from 623 studies published between 1977-1987 in the journal \emph{Ecology}. The mean duration of the most comparable subset of those values (n=419) was 3.6 years, versus 3.3 years in our sample (or 5.1 when excluding temporally un-replicated observations). The second dataset is found in Kareiva and Anderson \cite{kareiva_spatial_1988}, who assessed the resolutions of 97 community ecology experiments published in \emph{Ecology} between 1980-1986. The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,479,465 m$^2$), but comparing Kareiva and Anderson's \cite{kareiva_spatial_1988} 80th percentile value (197 m$^2$) to ours (115 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments. Porter et al \cite{porter_wireless_2005} provide the third dataset, which compares the extent and interval of 25 studies published in 2003 and 2004 (also in \emph{Ecology}). The mean interval was 178 days, compared to 684 days in our sample, but the 80th percentile value in our study was 169 days compared to 329 days in theirs. Extent in our sample was substantially larger according to multiple summary statistics, including the mean (368,403 ha versus 114,965,072 ha in our study), median (9 ha versus 5,051 ha), and 95th percentile (136,000 ha versus 46,424,808 ha; this value is smaller than the mean, which is skewed by a small number of very large extent observations).

Although limited due to methodological differences (e.g. a focus on experiments versus un-manipulated systems), these comparisons suggest that the duration and resolution of ecological observations have changed little in the past 30 years, but observational frequency and extent have both increased. A weak positive trend in our own data also suggests that the mean extent of ecological observations is steadily increasing (Supplementary Figure 5), which likely corresponds to increasing use of remote sensing (Supplementary Figure 4).

Despite this apparent increase in observational extent, there remains a large gulf between the areas that ecologists actually observe and the areas their observations are intended to represent (Fig. 3A). A substantial discrepancy also exists between the amount of time spent observing phenomena and the time spans those observations theoretically represent (Fig. 3B). These differences between the actual and ostensible scales of observation have implications for ecological understanding, as the unobserved portions of space and time may contain important patterns and processes that are not captured by replicates, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997,palmer_scale_1994, cao_comparison_2002, legendre_spatial_1993,collins_method_2000}. Brief, infrequent snapshots, or fine-grained, spatially sparse replicates, may be sufficient to characterize many phenomena (e.g. annual changes in tree cover are well-represented by low frequency satellite imaging \cite{hansen_high-resolution_2013}), but may be inadequate for more dynamic phenomena. For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to observe fire behavior \cite{clements_observing_2007}. To capture such behavior, long periods of continuous observation may be more important for understanding dynamics than frequent repeats.

It is therefore important to examine whether the scales of the phenomena being observed are adequately captured by the design of replicates. Our methods suggest one possible procedure for assessing the \emph{scale representativeness} of observations, which is to 1) calculate the autocorrelation (spatial or temporal) within the observations (using a semi-variogram, for example), 2) find the threshold distance (or time) below which a suitably strong correlation (e.g. r=0.7) will exist between neighboring sampled values, 3) add that distance (or time) to the sample resolution (or duration), and 4) recalculate actual extent (or duration) using the adjusted resolution (or sampling duration). The difference between this autocorrelation-adjusted actual extent (or actual duration) and extent (or duration) may provide a useful additional measure of how well the replicates represent the intended scale of observation. Though increasing spatial or temporal coverage may not always be the goal of a study, if the gap between actual and ostensible values remains large, then alternative sampling methods may be used to close it. For example, remote sensing provides wall-to-wall spatial coverage of a study area, erasing the difference between actual extent and extent. Furthermore, the interval of high-resolution imaging (higher resolution is preferred in images as it allows individual features to be better discerned \cite{dark_modifiable_2007, hay_comparison_2003}) is now approaching daily to sub-daily scales \cite{drusch_sentinel-2:_2012, hand_startup_2015}, allowing improved representation of spatial and temporal dynamics. For phenomena that can't be measured from space, either because they are not visible or because they require continuous observation, new approaches for collecting \emph{in situ} or near-surface observations (e.g. low-cost wireless sensors \cite{wolf_gsm-based_2012, collins_new_2006, porter_wireless_2005}, citizen observers \cite{dickinson_current_2012}, and autonomous vehicles \cite{anderson_lightweight_2013}) can be used to increase the spatial and temporal coverage of observations.

The aforementioned insights regarding modern observational domains must be tempered by the uncertainty within our own scale estimates, as detailed in the preceding section. However, most of this uncertainty is attributable to unclear reporting of scale values in the majority of papers we reviewed (a problem also noted in geography studies \cite{margulies_ambiguous_2016}). This tendency towards vague documentation offers one final insight: despite decades of accumulated knowledge regarding its importance \cite{levin_problem_1992, wiens_spatial_1989, chave_problem_2013, wheatley_factors_2009}, scale appears to remain a low priority throughout much of the ecological discipline. Beyond contributing to the broader problem of scientific reproducibility \cite{baker_1500_2016}, inattentiveness to scale increases the risk that observations inadequately represent the phenomenon of interest, thereby limiting the generalizability of any derived ecological knowledge \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989}. To mitigate this problem, we recommend that ecological journals require authors to quantify and clearly report the values of resolution, extent, interval, and duration. Fortunately, some journals already appear to be implementing such policies; for example, \emph{Global Ecology and Biogeography} now requires information on the spatial, temporal and taxonomic scale of studies to be in the abstract (a policy adopted in early 2016).

>>>>

%\linenumbers
\vspace{10pt}
\noindent \textbf{Introduction}
\vspace{5pt}
\\
The scales at which ecosystems are observed play a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.  Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is essential for understanding ecological phenomena \cite{levin_problem_1992,sandel_scale_2009}. This awareness has grown rapidly since the 1980s\cite{schneider_rise_2001}, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad and increasing range of scales \cite{schneider_rise_2001, chave_problem_2013}.

Given the growing awareness of scale, expanding data gathering capabilities, and the fact that the most comprehensive (and arguably best-known) meta-analyses \cite{tilman_ecological_1989,kareiva_spatial_1988} of ecological research scales were published nearly 30 years ago (but see \cite{porter_wireless_2005, sandel_scale_2009} for more recent reviews), it is both timely and important to assess the scales of contemporary ecological investigation. To address this need, we quantified the spatial and temporal domains of empirical observations that were reported within recently (2004-2014) published ecological studies. We define domain as the distribution of observations within the spectrum of one or more scale dimensions\footnote{This definition differs slightly from Wiens' \cite{wiens_spatial_1989}, who defined ``domain of scale'' as ``a portion of the scale spectrum within which process-pattern relationships are consistent regardless of scale.''}, and empirical observations as ecological observations collected under un-controlled or non-manipulated conditions. Empirical observations are critical for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992, tilman_ecological_1989}, therefore the spatio-temporal domains of observations provide an important indicator of the field's progress towards achieving a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.

@@

\vspace{10pt}
\noindent \textbf{Results}
\vspace{5pt}
\\
We reviewed 348 papers randomly selected from 42,918 titles published between 2004-2014 in the top 30 ecology-themed journals. We extracted scale data from 378 observations of ``natural'' (i.e. non-experimentally manipulated) ecological features reported within 133 of the reviewed papers (plus an additional 62 cited as the source of observations). Most sampled observations were collected using conventional field methods (80\%), followed by automated \emph{in situ} sensing techniques (12.4\%), remote sensing (6.9\%), and paleo-reconstruction ($<$0.8\%).

\vspace{5pt}
\noindent \textbf{\emph{Observational domains within individual dimensions}}\\
In terms of resolution, spatial replicates for most (67\%) observations (across all methods) were $\leq$1 m$^2$, a further 24\% were 1 m$^2$ up to 1 ha, and the remaining 9\% were $>$1 ha (Fig. 1A). These distributions primarily reflect those of field observations, the dominant observational methodology. Automated sensing and paleo-reconstruction observations had resolutions that were generally finer (85\% or more $\leq$0.1 m$^2$) than field observations (47\% $\leq$0.1 m$^2$), while most remote observations were much coarser (70\% $>$100 m$^2$; Supplementary Figure 1).

The extent of 19\% of observations was $\leq$10 ha, 23\% covered 10-1,000 ha, 11\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 15\% $>$1,000,000 ha (Fig. 1B).  As with resolution, the extent covered by automated sensing methods tended to be smaller (52\% $\leq$100 ha) than those of field observations (31\% $<$100 ha), while 96\% of remote and all paleo-reconstruction observations covered areas $>$10,000 ha.

@@

Duration was $\leq$1 day for 31\% of sampled observations (due to lack of temporal replication), while 10\% covered one day to one month, 23\% one month to one year, 27\% 1-10 years, and 9\% $>$1 decade (Fig. 1D). Paleo-reconstructions naturally had the longest duration (67\% $>$ 1 decade), while just $\sim$40\% of field, automated, and remote observations had durations exceeding one year.

\vspace{5pt}
\noindent \textbf{\emph{Observational domains within two-dimensions}}\\
Contrasting resolution with interval reveals that most temporally replicated observations had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig. 2A). A less dense, oblong concentration of observations bounded on the upper left by monthly to yearly observations at 100 m$^2$ resolution and on the lower right by near-daily to monthly observations with 1-10 ha resolution is also evident. The four observational methods occupied substantially different domains, as indicated by the locations of their median values (and see Supplementary Figure 2): the median domain of field observations had 0.1-1 m$^2$ resolution and a monthly interval, whereas remote observations had a coarser median resolution (1,000 m$^2$) but finer median interval ($\sim$1 day). Paleo-reconstructions and automated sensing were both finely resolved (median between 10 cm$^2$-0.01 m$^2$), but automated approaches had an hourly-daily median interval compared to multi-decadal for paleo-reconstructions.

Comparing the interval and duration of temporally replicated observations showed most observations had daily to decadal intervals and durations of one month to one decade (Fig. 2B). Interval appears to increases with duration; observations lasting one month to one year tend to have daily to monthly intervals, while those lasting one year to one decade tend to have yearly to decadal intervals. This tendency is reflected in the domain medians of the primary observational methods: automated sensing had the finest median interval (hour-day) and shortest duration (month-year), followed by remote sensing (median interval $\sim$1 day and duration 1 year), field observations (median interval 1 month and duration $\sim$1 year), and finally paleo-reconstructions (median interval 1 decade and millennial duration).

Contrasting the two spatial dimensions against one another shows a primary concentration of observations of 10 cm$^2$ to nearly 100 m$^2$ resolution with extents ranging between $\sim$1,000-1,000,000 ha (Fig. 2C). Another prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third, fainter concentration of 1-1,000 cm$^2$ resolution and 1000 m$^2$ to $\leq$10 ha extent. These three concentrations suggest that observational extent increases with resolution, which is further evident in the median domain values (and kernel densities; Supplementary Figure 2) of automated (0.01 m$^2$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.  Paleo-reconstructions were the exception, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a possible artifact of small sample size.

There are two primary observational domains within the contrast between duration and extent. The first consists of observations lasting one month to one decade with extents of 10-1,000 ha, while the second is defined by observations of one year to several decades that cover 10,000-1,000,000 ha (Fig. 2D).  Three other notable, but lesser concentrations are also evident, including small area observations (0.1-1 ha) covering one month to decade, and short duration, temporally unreplicated observations ($\leq$1 day) of either 1-100 ha or 10,000-1,000,000 ha.  The median observation from automated sensing (1 year duration, 100 ha extent) lies near the center of the first major concentration, while the median extents of field (1,000-10,000 ha) and remote (1-10 million ha) observations bound the second major concentration at its upper and lower extents, with the median duration of both observational types falling between one month to one year.

\vspace{5pt}
\noindent \textbf{\emph{Differences between actual and ostensible scales}}\\
Observational extent was on average 5.6 orders of magnitude larger than actual extent (Fig. 3A). This difference increased with extent, reaching a maximum of 8.3 between 100 million and 1 billion ha of extent, then falling to 3 orders of magnitude between 1-10 billion ha (these extents comrpised $<$2\% of observations, which were primarily collected with remote sensing). Remote observations had the smallest mean difference magnitude (1.9), compared to $\geq$5.7 for the other three methods (Supplementary Figure 3).

The difference magnitudes between observational duration and actual duration were somewhat smaller, averaging 3.4 and ranging from $\sim$2 for the shortest durations (hour-day) to >4 for observations lasting one decade to one century (Fig. 3B). As with extent, the difference fell substantially for the longest durations (century to millenia), as these domains were covered by paleo-reconstructions (Supplementary Figure 3), which have little difference between actual and ostensible duration because coring techniques capture continuous temporal records. The mean difference magnitudes for the other three observational methods ranged from just over 3 (field and automated sensing) to nearly 6 (remote observations).

\vspace{5pt}
\noindent \textbf{\emph{Potential biases and uncertainties in quantifying scales}}\\
Our results were potentially influenced by several methodological issues. First, most studies did not precisely report observational scales, thus we had to estimate, rather than simply record, scale values for most observations (we estimated 63, 60, 69, 36, 64, and 83\% of resolution, extent, actual extent, interval, duration, and actual duration values, respectively). Estimation errors may therefore have biased our findings. We attempted to quantify and account for this error by assessing between-observer variability and incorporating this uncertainty into our resampling methodology (Supplementary Results). The resulting confidence intervals (Fig. 1) suggest that estimation errors did not unduly influence our findings.

Our scale-estimation protocols may also have introduced bias, particularly our protocol for estimating resolution (the smallest areal unit of \emph{complete} measurement). We selected this definition for the sake of consistency, but some papers reported resolution as a larger area in which sub-samples were taken. For these, our estimates were finer than what the studies' authors considered to be the resolution. Our results would also  be somewhat different if we had included observations from experiments. For example, average resolution and duration would likely be finer \cite{tilman_ecological_1989,kareiva_spatial_1988}. Additionally, the token one second (Supplemental Methods) we used to represent the duration of remotely sensed temporal replicates (which are effectively instantaneous) caused us to underestimate the differences between their durations and actual durations (Supplementary Figure 3). However, the relatively small number of remote observations suggests that the impact of this bias on our overall findings was negligible.

@@

Finally, our omission of papers published after 2014 could also have biased our findings. Although our sample size was too small to assign statistical significance, we found a possible positive trend in the use of remote observations and a corresponding decline in field observations over the course of our study period. If these trends were not spurious, they suggest that including studies from 2015-2017 would result in a somewhat larger sample of remote observations, which could slightly increase the mean observational extent (see Supplementary Results).

\vspace{10pt}
\noindent \textbf{Discussion}
\vspace{5pt}
\\
\noindent \textbf{\emph{Insights into the scale domains of modern ecology}}\\
Our results suggest that modern ecology's observational domains are fairly narrow, and that ecologists still primarily rely on conventional field-based observational techniques. In the spatial dimensions, most observations have resolutions $\leq$1 m$^2$ and extents $\leq$10,000 ha (Fig 1A;B). In the temporal dimensions, most observations are either un-replicated or relatively infrequent ($>$1 month interval; Fig. 1C) and have relatively short duration ($\leq$1 year; Fig. 1D).

Contrasting observational dimensions reveals that larger extents are associated with larger spatial replicates (Fig. 2C), while longer durations are associated with longer intervals (Fig. 2B). The latter association reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of field observations, but also appears to affect the other three methods, as evidenced by their relative domain locations. A similar tradeoff is illustrated by the inverse relationship between resolution and interval (Fig. 2A), which primarily relates to field observations, where larger spatial replicates demand greater effort, reducing sampling frequency \cite{kareiva_spatial_1988}. Less obvious is the opposite tradeoff that affects remote observation (Supplementary Figure 2), where finer resolution (necessary for detail) typically necessitates longer intervals \cite{estes_platform_2016}.

As a result of these tradeoffs, there are several notable observational gaps, specifically within the domains defined by high frequency (daily to sub-daily intervals) observations having high to moderate resolutions ($>$1 m$^2$ up to 100 ha; Fig. 2A) and decadal or longer durations (Fig. 2B).  Another gap is evident in the high to moderate resolution, large extent (1 million-10 billion ha) domain (Fig. 2C).

Have these domains changed since the seminal papers on scale first appeared in the late 1980s \cite{wiens_spatial_1989, levin_problem_1992, tilman_ecological_1989}? A comprehensive answer would require a similar analysis focused on earlier literature, but the data provided by three prior studies provide partial insight.  The first dataset consists of duration values that Tilman \cite{tilman_ecological_1989} extracted from 623 studies published between 1977-1987 in \emph{Ecology}. The mean duration of the most comparable subset of those values (n=419) was 3.6 years, versus 3.3 years in our sample (or 5.1 when excluding temporally un-replicated observations). The second dataset is found in Kareiva and Anderson \cite{kareiva_spatial_1988}, who assessed the resolutions of 97 community ecology experiments published in \emph{Ecology} between 1980-1986. The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,479,465 m$^2$), but comparing Kareiva and Anderson's \cite{kareiva_spatial_1988} 80th percentile value (197 m$^2$) to ours (115 m$^2$) shows that most contemporary observations are finer-grained than most 1980s-era experiments. Porter et al \cite{porter_wireless_2005} provide the third dataset, which compares the extent and interval of 25 studies published in 2003-2004 in \emph{Ecology}. The mean interval was 178 days, compared to 684 days in our sample, but the 80th percentile value in our study was 169 days compared to 329 days in theirs. Extent in our sample was substantially larger according to multiple summary statistics, including the mean (368,403 ha versus 114,965,072 ha in our study), median (9 ha versus 5,051 ha), and 95th percentile (136,000 ha versus 46,424,808 ha; this value is smaller than the mean, which is skewed by a small number of very large extent observations).

Although limited due to methodological differences (e.g. a focus on experiments versus un-manipulated systems), these comparisons suggest that the duration and resolution of ecological observations have changed little in the past 30 years, but observational frequency and extent have both increased. A weak positive trend in our data also suggests that the mean extent of ecological observations is steadily increasing (Supplementary Figure 5), which likely corresponds to increasing use of remote sensing (Supplementary Figure 4).

Despite this apparent increase in observational extent, there remains a large gulf between the areas that ecologists actually observe and the areas their observations are intended to represent (Fig. 3A). A substantial discrepancy also exists between the amount of time spent observing phenomena and the time spans those observations theoretically represent (Fig. 3B). These differences between the actual and ostensible scales of observation have implications for ecological understanding, as the unobserved portions of space and time may contain important patterns and processes that are not captured by replicates, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997,palmer_scale_1994, cao_comparison_2002, legendre_spatial_1993,collins_method_2000}. Brief, infrequent snapshots, or fine-grained, spatially sparse replicates, may be sufficient to characterize many phenomena (e.g. annual changes in tree cover are well-represented by low frequency satellite imaging \cite{hansen_high-resolution_2013}), but may be inadequate for more dynamic phenomena. For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that they cannot be used to observe fire behavior \cite{clements_observing_2007}. To capture such behavior, long periods of continuous observation may be more important  than frequent repeats for understanding dynamics.

It is therefore important to examine whether the scales of the phenomena being observed are adequately captured by the design of replicates. Our methods suggest one possible procedure for assessing the \emph{scale representativeness} of observations, which is to 1) calculate the autocorrelation (spatial or temporal) within the observations (using a semi-variogram, for example), 2) find the threshold distance (or time) below which a suitably strong correlation (e.g. r=0.7) will exist between neighboring sampled values, 3) add that distance (or time) to the sample resolution (or duration), and 4) recalculate actual extent (or duration) using the adjusted resolution (or sampling duration). The difference between this autocorrelation-adjusted actual extent (or actual duration) and extent (or duration) may provide a useful additional measure of how well the replicates represent the intended scale of observation. Though increasing spatial or temporal coverage may not always be the goal of a study, if the gap between actual and ostensible values remains large, then alternative sampling methods may be used to close it. For example, remote sensing provides wall-to-wall spatial coverage of a study area, erasing the difference between actual extent and extent. Furthermore, the interval of high-resolution imaging (higher resolution is preferred in images as it allows individual features to be better discerned \cite{dark_modifiable_2007, hay_comparison_2003}) is now approaching daily to sub-daily scales \cite{drusch_sentinel-2:_2012, hand_startup_2015}, allowing improved representation of spatial and temporal dynamics. For phenomena that cannot be measured from space, either because they are not visible or because they require continuous observation, new approaches for collecting \emph{in situ} or near-surface observations (e.g. low-cost wireless sensors \cite{wolf_gsm-based_2012, collins_new_2006, porter_wireless_2005}, citizen observers \cite{dickinson_current_2012}, and autonomous vehicles \cite{anderson_lightweight_2013}) can be used to increase the spatial and temporal coverage of observations.

The aforementioned insights regarding modern observational domains must be tempered by the uncertainty within our own scale estimates, as detailed in the preceding section. However, most of this uncertainty is attributable to unclear reporting of scale values in the majority of papers we reviewed (a problem also noted in geography studies \cite{margulies_ambiguous_2016}). This tendency towards vague documentation offers one final insight: despite decades of accumulated knowledge regarding its importance \cite{levin_problem_1992, wiens_spatial_1989, chave_problem_2013, wheatley_factors_2009}, scale appears to remain a low priority throughout much of the ecological discipline. Beyond contributing to the broader problem of scientific reproducibility \cite{baker_1500_2016}, inattentiveness to scale increases the risk that observations inadequately represent the phenomenon of interest, thereby limiting the generalizability of any derived ecological knowledge \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989}. To mitigate this problem, we recommend that ecological journals require authors to quantify and clearly report the values of resolution, extent, interval, and duration. Fortunately, some journals already appear to be implementing such policies. For example, \emph{Global Ecology and Biogeography} now requires information on the spatial, temporal and taxonomic scale of studies to be in the abstract (a policy adopted in early 2016).
